### 2023-07-12 rapport/proglin.tex

\section{Résolution de la programmation linéaire \texorpdfstring{p}{$p$}-adique}

\begin{algorithm}
\caption{Un joli algorithme}
\begin{algorithmic}
\REQUIRE{habiter près des montagnes}
\REPEAT
 \IF{il fait beau}
  \STATE faire une randonnée
 \ELSE[il fait moche]
  \STATE résoudre P $\neq$ NP
 \ENDIF
\UNTIL{foulure de cheville}
\ENSURE{bobo}
\end{algorithmic}
\end{algorithm}