### 2025-12-09 issac2026/main.tex
commit by Simone

\documentclass[sigconf]{acmart}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs} 
\usepackage{graphicx}
\usepackage{color,xcolor}
%\usepackage{ulem}
\usepackage{bm}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}

\usepackage{hyperref}

\usepackage{cleveref}

% --------- ALGORITHMS ---------- <<:
\usepackage{algorithm,algorithmic}
%\usepackage[noend]{algpseudocode}      %%% this is for the algorithms
\renewcommand{\algorithmicrequire}{\textbf{Input:}} % require --> input
\renewcommand{\algorithmicensure}{\textbf{Output:}} % ensure --> output
% :>> ALGORITHMS

\usepackage{stmaryrd} %for double brackets

\newcommand{\R}{\mathbb{R}} % real numbers
\newcommand{\C}{\mathbb{C}} % complex numbers
\newcommand{\N}{\mathbb{N}} % integers
\newcommand{\allmat}{\mathbb{M}} % matrices
\newcommand{\sym}{\mathbb{S}} % matrices symétriques
\newcommand{\PP}{\mathcal{P}}
\newcommand{\Qc}{\mathcal{Q}}
\renewcommand{\span}[1]{{\text{span}(#1)}} % simone's comments
\newcommand{\aff}[1]{{\text{aff}(#1)}} % simone's comments
\newcommand{\relint}[1]{{\text{relint}(#1)}} % simone's comments
\newcommand{\calL}{\mathcal{L}} % simone's comments

\newcommand{\simone}[1]{{\color{blue} #1}} % simone's comments
\newcommand{\corentin}[1]{{\color{red} #1}} % corentin's comments
\newcommand{\tristan}[1]{{\color{red} #1}} % tristan's comments
\newcommand{\todo}[1]{{\color{blue}\bf \text{TODO: } #1}} % TODOs

%p-adic commands
\DeclareMathOperator{\val}{val}
\def\QQ{\ensuremath{\mathbb{Q}}}
\def\ZZ{\ensuremath{\mathbb{Z}}}
\newcommand{\OK}{\mathcal{O}_K}
\def\diag{\mathrm{diag}}

\newcommand{\GL}{\mathrm{GL}}

\newcommand{\exend}{\hfill $\blacksquare$}

%\usepackage{biblatex}
%\addbibresource{../rapport/bibstage.bib}

\title{On semidefinite-representable sets over valued fields}

\author{Simone Naldi}
\affiliation{
    \institution{Université de Limoges, CNRS, XLIM}
    \city{Limoges}
    \country{France}
}

\author{Tristan Vaccon}
\affiliation{
    \institution{Université de Limoges, CNRS, XLIM}
    \city{Limoges}
    \country{France}
}

\author{Corentin Cornou}
\affiliation{
    \institution{}
    \city{}
    \country{France}
}

\copyrightyear{2026}
\acmYear{2026}
\setcopyright{cc}
\setcctype{by}
\acmConference[ISSAC '26]{International Symposium on Symbolic and Algebraic Computation}{Date}{Place}
\acmBooktitle{International Symposium on Symbolic and Algebraic Computation (ISSAC '26), date, place}\acmDOI{}
\acmISBN{}

%\begin{CCSXML}
%<ccs2012>
%   <concept>
%       <concept_id>10010147.10010148.10010149.10010150</concept_id>
%       <concept_desc>Computing methodologies~Algebraic algorithms</concept_desc>
%       <concept_significance>500</concept_significance>
%       </concept>
% </ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Computing methodologies~Algebraic algorithms}

\keywords{keywords}

\begin{abstract}
  Real polyhedra and real spectrahedra (or their projections) are respectively the feasible sets
  of linear and semidefinite programming, and form the family of
  semidefinite-representable sets. This paper investigates versions of these sets and of
  the related optimization problems with data over a valued field $K$.
  For $K$-polyhedra and linear programming over $K$ we give an algorithm based on computation
  of Smith normal forms.
  We prove that basic properties of semidefinite-representable sets generalize to this setting.
  In particular, we showcase examples of non-polyhedral $K$-spectrahedra and of sets that are
  semidefinite representable over $K$ but are not $K$-spectrahedra.
\end{abstract}

\begin{document}
\maketitle

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}